### 2014-04-07 ApTranscriptome_TR.Rmd
Working on manuscript

**March 31, 2014**

@@

# apply `transcriptSD` function to all transcripts
A22.int.sd <- unlist(Map(transcriptSD, A22.int.lm, colony = "A22"))

# repeat for Ar
Ar.int.lm <- RxNlmAIC[Ar.int.transcripts$Transcript]
Ar.int.sd <- unlist(Map(transcriptSD, Ar.int.lm, colony = "Ar"))

# T-test comparing standard deviation of expression between colonies
t.test(A22.int.sd, Ar.int.sd)
```

@@

```{r plotthermalbreadth, echo=FALSE, eval=TRUE)
# prep data for plotting
A22.int.sd <- data.frame(colony = "A22", exp_sd = A22.int.sd)
Ar.int.sd <- data.frame(colony = "Ar", exp_sd = Ar.int.sd)
Ap.int.sd <- rbind(A22.int.sd, Ar.int.sd)

@@

```{r offset, echo=TRUE, eval=TRUE}
testset <- responsive.transcripts[1:10]
varshape.out <- ldply(responsive.transcripts, .progress = "text", RxNvarshape)

# to compare among transcripts, standardize by overall mean for each transcript

# t test
t.test(varshape.out$A22.mean, varshape.out$Ar.mean)
t.test(varshape.out$A22.slope, varshape.out$Ar.slope)
t.test(varshape.out$A22.curve, varshape.out$Ar.curve)
t.test(varshape.out$A22.wiggle, varshape.out$Ar.wiggle)

>>>>

**April 4, 2014**

@@

# apply `transcriptSD` function to all transcripts
A22.int.sd <- unlist(Map(transcriptSD, A22.int.lm, colony = "A22"))
A22.int.sd <- data.frame(colony = "A22", exp_sd = A22.int.sd)

# repeat for Ar
Ar.int.lm <- RxNlmAIC[Ar.int.transcripts$Transcript]
Ar.int.sd <- unlist(Map(transcriptSD, Ar.int.lm, colony = "Ar"))
Ar.int.sd <- data.frame(colony = "Ar", exp_sd = Ar.int.sd)

# T-test comparing standard deviation of expression between colonies
t.test(A22.int.sd$exp_sd, Ar.int.sd$exp_sd)
```

@@

```{r plotthermalbreadth, echo=FALSE, eval=TRUE)
# prep data for plotting
Ap.int.sd <- rbind(A22.int.sd, Ar.int.sd)

@@

```{r offset, echo=TRUE, eval=TRUE}
varshape.out <- ldply(interaction.transcripts, .progress = "text", RxNvarshape)

# plot
boxplot(varshape.out$A22.mean, varshape.out$Ar.mean)
boxplot(log(varshape.out$A22.mean), log(varshape.out$Ar.mean))

# calculate differences in mean, slope, curvature and wiggle between colonies for each transcript
# offset = mean1 - mean2
varshape.out$offset <- varshape.out$A22.mean - varshape.out$Ar.mean
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.offset <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.offset"] <- (varshape.out[i, "offset"] / mean(c(varshape.out[i, "A22.mean"], varshape.out[i, "Ar.mean"])))
}

# t-test
t.test(varshape.out$s.offset, alternative = "two.sided")

# slope = slope1 - slope2
varshape.out$slope <- varshape.out$A22.slope - varshape.out$Ar.slope
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.slope <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.slope"] <- (varshape.out[i, "slope"] / mean(c(varshape.out[i, "A22.slope"], varshape.out[i, "Ar.slope"])))
}
# for transcripts where both samples had slope=0, change s.slope from Inf to 0 as there is truly no difference in slope
varshape.out[which(varshape.out$Ar.slope == 0 & varshape.out$A22.slope == 0), "s.slope"] <- 0

# t-test
t.test(varshape.out$s.slope, alternative = "two.sided")

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
t.test(varshape.out$s.curvature, alternative = "two.sided")

# wiggle
varshape.out$wiggle <- varshape.out$A22.wiggle - varshape.out$Ar.wiggle
varshape.out$s.wiggle <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.wiggle"] <- (varshape.out[i, "wiggle"] / mean(c(varshape.out[i, "A22.wiggle"], varshape.out[i, "Ar.wiggle"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.wiggle == 0 & varshape.out$A22.wiggle == 0), "s.wiggle"] <- 0
# note that only 407 transcripts have any 'wiggle'
length(which(varshape.out$s.wiggle != 0))

# t-test
t.test(varshape.out$s.wiggle, alternative = "two.sided")
```

Next, I partition the differences in the reaction norms into the variation explained by differences in the trait means (offset), slope, curvature and wiggle.

```{r partition_reaction_norm_variation}
# take absolute value of each value and sum to get total differences between reaction norms
varshape.out$s.total <- abs(varshape.out$s.offset) + abs(varshape.out$s.slope) + abs(varshape.out$s.curvature) + abs(varshape.out$s.wiggle)

t.test(varshape.out$s.total, alternative = "two.sided")

# variation in reaction norms due to differences in mean
varshape.out$prop.mean <- abs(varshape.out$s.offset) / varshape.out$s.total
varshape.out$prop.slope <- abs(varshape.out$s.slope) / varshape.out$s.total
varshape.out$prop.curve <- abs(varshape.out$s.curvature) / varshape.out$s.total
varshape.out$prop.wiggle <- abs(varshape.out$s.wiggle) / varshape.out$s.total

mean(varshape.out$prop.mean)
mean(varshape.out$prop.slope)
mean(varshape.out$prop.curve)
mean(varshape.out$prop.wiggle)
```
